Online {\IterComp}

### 2017-09-16 related.tex
done with related work

\subsection{{\IterComp}}
Various adaptive compilation techniques have been proposed for finding near optimal compilation
settings~\cite{agakov06,kulkarni04,stephenson03}. These approaches evaluate the performance of candidate compiler options through repeated
executions using the same input. All these approaches implicitly assume the best-performing compiler settings found from a small set of
inputs will work well on unseen datasets. Chen~\etal~\cite{chen10,chen12a} show that the best optimization options can differ across
inputs. Their results suggest that, in most cases it is possible to find a combination of optimizations that produces good results across
all inputs, by using a few tens of inputs to evaluate our optimizations on. While a positive result, this is still impractical. Producing
tens of realistic input sets might be difficult. Even if not, each combination of optimization options has to be evaluated on every input
set, increasing the time \itercomp takes by at least an order of magnitude.
%=======
%Multiple adaptive compilation techniques have been proposed for finding a near optimal combination of optimization
%settings~\cite{agakov06,hoste08,kulkarni04,stephenson03} by probabilistically searching the space of possible combinations. But with
%all of them there is uncertainty about whether the quality of their results depends on the data set used. Chen~\etal~\cite{chen10,chen12a}
%examined this issue and showed that indeed the best optimization options differ across inputs. Still, in most cases it is possible to find
%an optimization sequence that produces good results across all inputs, by evaluating the optimizations on a representative set inputs.
%While a positive result, this is still impractical. Producing a set of realistic inputs might be difficult. Even if not, each
%optimization sequence has to be evaluated on every input, increasing the time \itercomp takes by at least an order of
%magnitude.
%>>>>>>> origin/master

Recent research~\cite{chen12b,fang15} has used real data for performing {\itercomp} on distributed data center applications. Each
worker receives a subset of the input dataset to evaluate a small set of optimization settings on. The best such setting from each round
is used for subsequent executions of the same code. By testing new settings and re-evaluating old ones on new datasets this approach makes
it more likely to select an optimization setting that works well across inputs. Like in~\cite{chen10} every setting is tested on the same
set of inputs, wasting time and energy. This approach only works well with MapReduce-like workloads, since it relies on the MapReduce
framework for repeating the same computation multiple times without side-effects or corrupting the input.

Fursin~\etal~\cite{fursin07} attempted to find a more general approach for online \itercomp. They proposed instructions per cycle (IPC) as
a metric for comparing runs of the same program with different optimization settings and different inputs: application binaries displaying
higher IPC are more efficient than lower IPC binaries. Their results show that IPC seems promising as a robust metric for {\itercomp}.
However, some optimization techniques affect IPC in the opposite way, making the program more efficient while lowering IPC. In the context
of multithreaded application, IPC cannot be used to estimate performance~\cite{alameldeen06,eyerman08}. Aalameldeen and
Wood~\cite{alameldeen06} suggest a work metric for quantifying efficiency, but choosing a unit of work is program specific and challenging.

@@

Donald Knuth~\cite{knuth73} introduced an optimal algorithm for profiling block frequency, inserting much fewer probes than na\"ive
instrumentation approaches that were used in practice~\cite{knuth71}. Forman~\cite{forman81} proposed further overhead reductions by
placing the probes in basic blocks that are less likely to be executed based on static heuristics. Ball and Larus~\cite{ball94} offered a
detailed discussion comparing two approaches for optimally profiling block frequency, namely, placing probes on the vertices or the edges
of the control-flow graph (CFG). They show that the edge-based approach produces optimal placement of probes. Theirs is the algorithm we
used for our initial selection of probes before relaxation.

>>>>

\subsection{{\IterComp}}
Various adaptive compilation techniques have been proposed for finding near optimal compilation
settings~\cite{agakov06,kulkarni04,stephenson03,hoste08}. These approaches compare the performance of candidate compiler options by first
compile the program with different options and then profiling the different compiled versions on the same program input. All these
approaches implicitly assume that the best-performing compiler settings found from a small set of inputs will work well on unseen datasets.
Chen~\etal~\cite{chen10,chen12a} show that the best optimization options can differ across inputs. Their results suggest that, in most
cases it is possible to find a single set of optimizations that produces good results across all inputs; but achieving this requires
developers provide a few tens of representative inputs to be used by the iterative compilation search process. We argue that providing
representative inputs is difficult in practice, as this require developers to foresee all possible user inputs. Even this is possible,
current approaches all require to evaluate each candidate compiler option on every single input. Doing so incurs significant overhead,
which thus prevents iterative compilation to be adopted at a larger scale.

%=======
%Multiple adaptive compilation techniques have been proposed for finding a near optimal combination of optimization
%settings~\cite{agakov06,hoste08,kulkarni04,stephenson03} by probabilistically searching the space of possible combinations. But with
%all of them there is uncertainty about whether the quality of their results depends on the data set used. Chen~\etal~\cite{chen10,chen12a}
%examined this issue and showed that indeed the best optimization options differ across inputs. Still, in most cases it is possible to find
%an optimization sequence that produces good results across all inputs, by evaluating the optimizations on a representative set inputs.
%While a positive result, this is still impractical. Producing a set of realistic inputs might be difficult. Even if not, each
%optimization sequence has to be evaluated on every input, increasing the time \itercomp takes by at least an order of
%magnitude.
%>>>>>>> origin/master

Recent studies~\cite{chen12b,fang15} have used real data to perform {\itercomp} on distributed data center applications. Each worker
receives a subset of the input dataset to evaluate a small set of optimization settings on. The best such setting from each round is used
for subsequent executions of the same code. By testing new settings and re-evaluating old ones on new datasets this approach makes it more
likely to select an optimization setting that works well across inputs. Like in~\cite{chen10} every setting is tested on the same set of
inputs, wasting time and energy. This approach only works well with MapReduce-like workloads, since it relies on the MapReduce framework
for repeating the same computation multiple times without side-effects or corrupting the input.

Fursin~\etal~\cite{fursin07} attempted to find a general approach for online \itercomp. They proposed to use instructions per cycle (IPC)
as a metric for comparing different runs of the same program with different optimization settings and different inputs. Their hypothesis is
that application binaries displaying higher IPC are more efficient than lower IPC binaries. However, many optimization techniques affect
IPC in the opposite way, making the program more efficient while lowering IPC. We have also shown in Section~\ref{sec:motivation}, IPC is
not strongly correlate to speedup. Therefore, it is not a good metric for \itercomp. Furthermore, IPC cannot be used to estimate
performance for multi-threaded applications~\cite{alameldeen06,eyerman08} due to the side-effect of thread contention. Aalameldeen and
Wood~\cite{alameldeen06} also suggest that relying on IPC can be harmful, and work metrics like time per work unit or work units per time
should be used. However, prior works all rely on developers to provide such metrics. This paper presents the first automatic approach to
derive such work metrics from arbitrary code without developer intervention.

@@

The seminal work of Donald Knuth~\cite{knuth73} introduced an optimal algorithm for profiling block frequency, inserting much fewer probes
than the na\"ive instrumentation approaches that were used in practice~\cite{knuth71}. Forman~\cite{forman81} proposed an approach to
further reduce the instrumentation overhead by placing probes in basic blocks that are less likely to be executed based on static
heuristics. Ball and Larus~\cite{ball94} offered a detailed discussion to compare these two approaches for optimally profiling block
frequency, namely, placing probes on the vertices or the edges of the control-flow graph (CFG). They show that the edge-based approach
produces better placement of probes. Our work built upon these prior works. We use the edge-based approach to place our probes before
applying relaxation.
